Clustering Beat-Chroma Patterns\\ in a Large Music Database

### 2010-03-23 Paper/Paper.tex
improved

\subsection{Definitions}
A pattern, or a patch, is a \textit{Shingle} taken from an actual song.
A code is similar to a pattern, but is the result of the clustering
algorithm and most likely does not come from an actual song.
Distortion is our measure of distance between a pattern and a code
(or between two patterns, see equation \ref{eq:dist} in Section 
\ref{sec:experiments}).

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.6\columnwidth]{code}
\end{center}
\caption{\small{A typical code from a codebook of size $200$. 
Patterns represented
$2$ bars and the pattern length was set to $8$.
}}
\label{fig:code}
\end{figure}

@@

\subsection{Codebook Learning}
We use an online version of the vector quantization algorithm 
\cite{Gersho1991}. This can also be seen as an online K-means.
The idea is as follow: take a sample from your data, find the closest
code in the codebook, bring that code closer to the sample by some amount.
The details are explained in Algorithm \ref{algo:vq}. Note that this
algorithm, though not optimal, is $O(n)$ with $n$ number of songs seen.
Thus it scales linearly to any size of data set. The setting of the learning
rate might become trickier, but this is not fundamentally a problem.

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.99\columnwidth]{data_sizes}
\end{center}
\caption{\small{Distortion for a codebook of size $100$ encoding one bar
at a time, a bar being represented by $4$ chroma feature vector.
Therefore, a code or a pattern is of size $12$ x $4$ = $48$.
Distortion measure on the test data set. Data size tested range
from $0$ (just initialization) to $500$ thousand. Patterns were selected at
random from the dataset of approximately $3,7$ million patterns, except
for the dotted line (non random in the legend). In that case we took the
first $N$ samples in our dataset. The difference is only visible for small
training sets ($1$K).
}}
\label{fig:data_sizes}
\end{figure}

@@

The 25 most frequently used codewords in the test set are shown in
Figure \ref{fig:codes1}.  
The frequency of use of these codewords are shown in Figure \ref{fig:freqs}.
The codewords primarily consist of sustained
notes and simple chords.  Since they are designed to be key-invariant,
specific notes do not appear.  Instead the first 7 codewords
correspond to a single note sustained across two bars (codeword 1),
perfect fifth (codes 2 and 3) and fourth intervals (codes 4 7), and a
major and minor triads (code 6 and 5).  Many of the remaining
codewords correspond to common transitions from one chord to another,
\eg a transition from one major triad to another (\ie Cmaj
$\rightarrow$ Fmaj assuming that chroma 0 is C) in codes 8 and 10 and
the reverse transition in code (\ie Fmaj $\rightarrow$ Cmaj) in code
22.  Does this transition have a better name like ``relative x''?

@@

Crazy idea: chord recognition via quantization: transcribe codewords,
then just encode song to get chords. would this work?

@@

\begin{table}
\begin{center}
\begin{tabular}{c|c}
%\hline
offset & \% of times chosen \\ \hline
0 & $\mathbf{62.6}$\\
1 & $16.5$\\
2 & $09.4$\\
3 & $11.5$\\
\end{tabular}
\end{center}
\caption{\small{
Best offset according to our codebook. The real one is $0$.
}}
\label{tab:offset}
\end{table}

>>>>

\subsection{Definitions}
A pattern, or a patch, is a beat-chroma pattern taken from an actual song.
A codeword (sometimes just code) is similar to a pattern, but is the result 
of the clustering
algorithm and most likely does not come from an actual song.
Distortion is our measure of distance between a pattern and a code
(or between two patterns, see equation \ref{eq:dist} in Section 
\ref{sec:experiments}).

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.6\columnwidth]{code}
\end{center}
\caption{\small{A typical codeword from a codebook of size $200$. 
Patterns represented
$2$ bars and the pattern length was set to $8$.
}}
\label{fig:code}
\end{figure}

@@

\subsection{Codebook Learning}
We use an online version of the vector quantization algorithm 
\cite{Gersho1991}. This can also be seen as an online K-means.
The idea is as follow: take a sample from your data, find the closest
codeword in the codebook, bring that codeword closer to the 
sample by some amount.
The details are explained in Algorithm \ref{algo:vq}. Note that this
algorithm, though not optimal, is $O(n)$ with $n$ number of songs seen.
Thus it scales linearly to any size of data set. The setting of the learning
rate might become trickier, but this is not fundamentally a problem.

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.99\columnwidth]{data_sizes}
\end{center}
\caption{\small{Distortion for a codebook of size $100$ encoding one bar
at a time, a bar being represented by $4$ chroma feature vector.
Therefore, a codeword or a pattern is of size $12$ x $4$ = $48$.
Distortion measure on the test data set. Data size tested range
from $0$ (just initialization) to $500$ thousand. Patterns were selected at
random from the dataset of approximately $3,7$ million patterns, except
for the dotted line (non random in the legend). In that case we took the
first $N$ samples in our dataset. The difference is only visible for small
training sets ($1$K).
}}
\label{fig:data_sizes}
\end{figure}

@@

The 25 most frequently used codewords in the test set are shown in
Figure \ref{fig:codes1}.  
The frequency of use of these codewords are shown in Figure \ref{fig:freqs}.
The codewords primarily consist of sustained
notes and simple chords.  Since they are designed to be key-invariant,
specific notes do not appear.  Instead the first 7 codewords
correspond to a single note sustained across two bars (codeword 1),
perfect fifth (codes 2 and 3) and fourth intervals (codes 4 7), and a
major and minor triads (codeword 6 and 5).  Many of the remaining
codewords correspond to common transitions from one chord to another,
\eg a transition from one major triad to another (\ie Cmaj
$\rightarrow$ Fmaj assuming that chroma 0 is C) in codes 8 and 10 and
the reverse transition in code (\ie Fmaj $\rightarrow$ Cmaj) in code
22.  Does this transition have a better name like ``relative x''?

@@

@@

\begin{table}
\begin{center}
\begin{tabular}{c|c}
%\hline
offset & \% of times chosen \\ \hline
0 & $\mathbf{62.6}$\\
1 & $16.5$\\
2 & $09.4$\\
3 & $11.5$\\
\end{tabular}
\end{center}
\caption{\small{
Best offset according to our codebook. The real one is $0$.
Random guess gives an accuracy of $25$\%.
}}
\label{tab:offset}
\end{table}
